---
title: 'Occupancy Sub Model C2: Core Home Range Scale Covariates'
author: "Anna Kurtin"
date: "2024-07-26"
output: html_document
---

This is the second iteration of the core home range (130 m radius of points) scale submodel within the occupancy/state submodel. From the first iteration, we saw that veg sd residual was hugely important in the model. It is also colinear with canopy height, canopy percent cover, and subcanopy height. Here, I am trying to break apart these colinearities to determine if the unique contribution of these coviariates is important of if their importance was only driven by total veg sd in the first model. I am also breaking apart the colinearity between subcanopy sd and subcanopy height. Subcanopy percent cover wasn't correlated with anything.

Even though there is not a causal relationship to the colinearity between total veg sd and percent canopy cover, there is a correlation between them that still needs to be broken up in the model. By running a regression and using the residual values of percent canopy cover, we aren't saying that one causes the other, we just say they are correlated and are pulling out the residual variation to isolate it's unique effect on the response. 

I think that by doing this, the covariates for canopy height, subcanopy height, and percent canopy may change in importance. 

**Model structure:**

State model with six covariates: percent canopy residuals from total veg sd (cov1) and subcanopy percent canopy (unchanged) (cov2), height of canopy residuals from total veg sd (cov3) and height of subcanopy residuals from subcanopy sd (cov4), standard deviation of subcanopy (unchanged) (cov5), total veg standard deviation (unchanged) (cov6)

$$
z_i \sim Bernoulli(\psi | \alpha_{1}, \alpha_{2}, \alpha_{3}, \alpha_{4}, \alpha_{5}, \alpha_{6})
$$ 

Naive process model:  

$$
y_{ij} \sim Bernoulli(p)
$$

**Handling of NAs:** omission for detections, imputation for habitat covariates

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(rjags)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(beepr)
library(corrgram)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
```

```{r Read in Data}
full_dat <- read.csv("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Habitat_Model_Covariates/Occupancy_Covariates/DetHist_VegCovs_Scaled_7-24.csv")

detections <- full_dat[,2:7]

# Have to specify the z data for a true presence if there was a 1 at any time 
z_dat <- rowSums(detections, na.rm = T)
z_dat[z_dat > 1] <- 1
z_dat[z_dat == 0] <- NA

# get the length of non-nas in the detection data
get_na <- function(x){(max(which(!is.na(x))))}
# Apply this to our data
miss <- apply(detections, 1, get_na)
```

```{R Format New Residual Covariates, message = FALSE}
# formula = y ~ x, data = 
# I did this on scaled covariates (no difference doing this on scaled vs unscaled data)

# Create covariates of residuals
# Canopy height from total veg st dev
reg1 <- glm(formula = ht_can_core ~ sd_allveg_core, data = full_dat)
cht_predict <- predict(reg1, full_dat)
cht_resid <- full_dat$ht_can_core - cht_predict
full_dat$canopy_ht_resid <- round(cht_resid,2)

# Subcanopy height from subcan st dev
reg2 <- glm(formula = ht_subcan_core ~ sd_subcan_core, data = full_dat)
scht_predict <- predict(reg2, full_dat)
scht_resid <- full_dat$ht_subcan_core - scht_predict
full_dat$subcan_ht_resid <- round(scht_resid,2)

# Canopy percent cover resid on total veg sd
reg3 <- glm(formula = pct_can_core ~ sd_allveg_core, data = full_dat)
cpct_predict <- predict(reg3, full_dat)
cpct_resid <- full_dat$pct_can_core - cpct_predict
full_dat$canopy_pct_resid <- round(cpct_resid,2)


modc2_dat <- full_dat %>% select(sd_allveg_core, 
                                 sd_subcan_core, 
                                 canopy_ht_resid, 
                                 subcan_ht_resid, 
                                 canopy_pct_resid,
                                 pct_subcan_core)
# test correlation
corrgram(modc2_dat, order=TRUE, lower.panel=panel.cor,
         upper.panel=panel.pts, text.panel=panel.txt,
         main="Correlations in Core Scale Metrics Model 2")

```

Now we can see that none of the new covariates are colinear with each other.

Let's run the model and see what changes!

```{R Setup JAGS Data}
# Create a list of data for JAGS
# index this by variable names to make sure nothing is being funky 
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z_dat,
  # Detection/nondetection data
  det_data = detections,
  # percent canopy residual on total veg sd
  cov1 = full_dat$canopy_pct_resid,
  # percent subcanopy unchanged
  cov2 = full_dat$pct_subcan_core,
  # canopy height residual on total veg sd
  cov3 = full_dat$canopy_ht_resid,
  # subcanopy height residual on subcanopy veg sd
  cov4 = full_dat$subcan_ht_resid,
  # subcanopy sd unchanged
  cov5 = full_dat$sd_subcan_core,
  # veg sd unchanged
  cov6 = full_dat$sd_allveg_core,
  n_unit = nrow(detections),
  miss = miss
)

# Set initial values 
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    a1 = rnorm(1, 0, 1),
    a2 = rnorm(1, 0, 1),
    a3 = rnorm(1, 0 ,1),
    a4 = rnorm(1, 0, 1),
    a5 = rnorm(1, 0, 1),
    a6 = rnorm(1, 0 ,1),
    b0 = rnorm(1, 0, 1)
  )
}
```

```{R Sub Model C1 Structure}
cat("
  model{
        # Priors
    a0 ~ dlogis(0, 1)
    a1 ~ dlogis(0, 1)
    a2 ~ dlogis(0, 1)
    a3 ~ dlogis(0, 1)
    a4 ~ dlogis(0, 1)
    a5 ~ dlogis(0, 1)
    a6 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
  
    # Likelihood
    # Loop through each sampling unit
    for(i in 1:n_unit){
      # Impute missing habitat data
      cov1[i] ~ dnorm(0, 1)
      cov2[i] ~ dnorm(0, 1)
      cov3[i] ~ dnorm(0, 1)
      cov4[i] ~ dnorm(0, 1)
      cov5[i] ~ dnorm(0, 1)
      cov6[i] ~ dnorm(0, 1)
  
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + a1*cov1[i] + a2*cov2[i] + a3*cov3[i] + a4*cov4[i] + a5*cov5[i] + a6*cov6[i]
      
      for(j in 1:miss[i]){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0
        
      }
    }
    
    # Derived parameters
    
  }
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/JAGS_HabOccModC2.txt")

```

```{R Run Model, results = FALSE}
fit_C2 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/JAGS_HabOccModC2.txt",
                 parameters.to.save = c("a0", "a1", "a2", "a3", "a4", "a5", "a6", "b0"),
                 n.iter = 10000, 
                 n.burnin = 4000, 
                 n.chains = 3,
                 n.thin = 2,
                 inits = init_func)
beep(sound = 3)
```

```{R Evaluate Model}
print(fit_C2)
tracedens_jags(fit_C2, parmfrow = c(3,3))
```

```{R Create Violin Plots}
chains_c2 <- jags_df(fit_C2)
# Select the chains for our covariates of interest
chains_viol <- chains_c2 %>% select(a1,a2,a3,a4,a5,a6)
# Rename them to be more interpretable
colnames(chains_viol) <- c("% canopy resid","% subcanopy","canopy height resid","subcanopy height resid", "subcanopy sd", "veg sd")
# Pivot this longer so that we can visualize it
chains_viol_long <- chains_viol %>% pivot_longer(cols = c("% canopy resid","% subcanopy","canopy height resid","subcanopy height resid", "subcanopy sd", "veg sd"),names_to = "Parameter", values_to = "Values")

# Create the violin plot
ggplot() +
  geom_violin(data=chains_viol_long, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Core Area Scale Covariates") +
  scale_fill_manual(values = c("% canopy resid"=palette_8[1], "% subcanopy" = palette_8[2],"canopy height resid"=palette_8[3], "subcanopy height resid" = palette_8[4], "subcanopy sd" = palette_8[5], "veg sd" = palette_8[6])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))
```



```{R Boo Spooky Graveyard, include = FALSE}
# testing out residuals on unscaled data to see if it's the same thing
# full_dat_t <- read.csv("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Habitat_Model_Covariates/Occupancy_Covariates/All_Veg_Covariates_7-24.csv")
# reg1_t <- glm(formula = ht_can_core ~ sd_allveg_core, data = full_dat_t)
# cht_predict_t <- predict(reg1, full_dat)
# cht_resid_t <- full_dat$ht_can_core - cht_predict_t
# test <- round(cht_resid_t,2)
# YES its the same thing
#full_dat_t$canopy_ht_resid_t <- round(cht_resid_t,2)
```