---
title: "Recover Parameters for Simulated Data Occupancy Model"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(coda)
library(janitor)
library(loo)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/6_Function_Scripts/Create_SiteColumn_fromPointID.R")
```


# Simulate Data


```{R generate data, out.width = "50%", out.height = "50%"}
# b1/cov1: tree richness
# b2/cov2: shrub community
# b3/cov3: subcanopy avg height
# b4/cov4: subcanopy percent cover
# b5/cov5: subcanopy veg complexity 
# b6/cov6: veg density 
# b7/cov7: canopy height
# b8/cov8: canopy percent cover

nP <- 107  # number of points
nV <- 6    # number of repeated visits
nC <- 8    # number of covariates
#survey_covs <- 3

# create latent state (z) and detection (y) data storage
z <- NULL        # point-level latent state
y <- array(NA, dim = c(nP, nV))   # point-level observation data

# create separate arrays for each covariate and fill with randomly generated data
# Covariate 1: Tree Richness
set.seed(22)
rich_vals <- c(-0.617, 0.71, 2.03, 3.37) # Create possible values for species richness
tree_rich <- sample(rich_vals, nP, replace =TRUE)
hist(tree_rich, col = cuckoo_palette[1], main = "Distribution of Simulated Native Broadleaf Richness")
# Expected response: linear and positive - higher levels of richness, higher insect abundance - more likely cuckoo presence 

# Covariate 2: Shrub Community
set.seed(3)
com_fs <- c(1:5)
shrub_com <- sample(com_fs, nP, replace = TRUE)
hist(shrub_com, col = cuckoo_palette[1], main = "Distribution of Simulated Shrub Community")
#shrub_com <- as.factor(shrub_com)
# Expected response: higher probability of occ with floodplain and broadleaf communities 

# Covariate 3: Subcanopy Veg Avg Height
set.seed(10)
subcan_h <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_h, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Height")
# Expected response: linear and positive - higher subcanopy veg means more subcanpy habitat and greater prob of occ

# Covariate 4: Subcanopy Veg Percent Cover
set.seed(11)
subcan_pc <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_pc, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Height")
# Expected response: linear and positive - higher subcanopy veg cover means more places to forage and hide and greater prob of occ

# Covariate 5: Subcanopy Veg Complexity 
set.seed(12)
subcan_cx <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_cx, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Complexity")
# Expected response: linear and positive - higher complexity means more places to forage and hide, greater prob of use 

# Covariate 6: Overall Veg Density
set.seed(13)
veg_d <- rnorm(nP, mean = 1, sd = 1)
hist(veg_d, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Total Vegetation Density")
# Expected response: linear and positive - higher veg cover means more places to forage and hide and greater prob of occ

# Covariate 7: Canopy Height
set.seed(14)
can_h <- rnorm(nP, mean = 1, sd = 1)
hist(can_h, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Canopy Height")
# Expected response: selecting for intermediate? want shrubby areas with some trees but not super matrure cottonwood gallery forest ???????????????????????????????????

# Covariate 8: Canopy Percent Cover
set.seed(15)
can_pc <- rnorm(nP, mean = 1, sd = 1)
hist(can_pc, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Canopy Cover")
# Expected response: also selecting for intermediate? want intermediate levels of canopy cover. since they prefer shrubby habitat. how to parameterize this ??????????????????????????????????? 
# BUT selecting for high canopy cover at a larger scale 

# covariate effects (pull randomly for each covariate)
set.seed(5)
beta <- rnorm(nC, 0, 1)

psi0 <- qlogis(0.2) # probability of presence
p0 <- qlogis(0.6) # probability of detection

psi <- NULL
for (i in 1:nP){
  # How do I account for categorical variables here? Just making them a factor
  psi[i] = plogis(psi0 + (beta[1]*tree_rich[i]) + (beta[2]*shrub_com[i]) + (beta[3]*subcan_h[i]) + (beta[4]*subcan_pc) + (beta[5]*subcan_cx) + (beta[6]*veg_d) + (beta[7]*can_h) + (beta[8]*can_pc))
}
# Assume equal probability of presence at all sites with this simulation

# get your true presence/non presence values from psi
set.seed(8) # I changed this around until I got a value of presence/absence that looked good to me
z <- rbinom(nP, 1, psi)

# total number of detections across number of visits (nV)
y <- matrix(NA, nP, nV)
# fill in the detection matrix dependent on whether the animal was there (z) and whether it was detected (p0)
# add covariate effects here
set.seed(7)
for (i in 1:nP){
  for (j in 1:nV){
    y[i,j] <- rbinom(1, z[i], plogis(p0))
    # Assume equal probability of detection at all sites and all surveys
  }
}

# b1/cov1: tree richness
# b2/cov2: shrub community
# b3/cov3: subcanopy avg height
# b4/cov4: subcanopy percent cover
# b5/cov5: subcanopy veg complexity 
# b6/cov6: veg density 
# b7/cov7: canopy height
# b8/cov8: canopy percent cover
```

```{R Set Up Model,eval = FALSE}
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_unit){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + (b1*tree_rich[i]) + (b2*shrub_com[i]) + (b3*subcan_h[i]) + (b4*subcan_pc[i]) + (b5*subcan_cx[i]) + (b6*veg_d[i]) + (b7*can_h[i]) + (b8*can_pc[i])
      
      cov2[i] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
      
      for(j in 1:n_rep){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0
        
      }
    }
    
    # Priors
    a0 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dnorm(0, 1)
    b2 ~ dnorm(0, 1)
    b3 ~ dnorm(0, 1)
    b4 ~ dnorm(0, 1)
    b5 ~ dnorm(0, 1)
    b6 ~ dnorm(0, 1)
    b7 ~ dnorm(0, 1)
    b8 ~ dnorm(0, 1)
  }
  ", file = "./R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel1.txt")
```

*Next up:* Figure out what the parameterization of these models means
```{r}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z,
  # Detection/nondetection data
  det_data = y,
  # Tree richness
  tree_rich = tree_rich,
  # Shrub community
  shrub_com = shrub_com,
  # Subcanopy height
  subcan_h = subcan_h,
  # Subcanopy cover
  subcan_pc = subcan_pc,
  # Subcanopy complexity 
  subcan_cx = subcan_cx,
  # Overall veg density
  veg_d = veg_d,
  # Canoy height
  can_h = can_h,
  # Canopy cover
  can_pc = can_pc,
  n_unit = nP,
  n_rep = nV
)

# Set initial values (optional)
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b2 = rnorm(1, 0, 1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1),
    b5 = rnorm(1, 0, 1),
    b6 = rnorm(1, 0, 1),
    b7 = rnorm(1, 0, 1),
    b8 = rnorm(1, 0, 1)
  )
}

ni <- 10000
nb <- 5000
nc <- 4
nt <- 2

# Run JAGS model on the same model we specify in the Fit_JAGS_Detection_Model2_HabChap markdown
occ_m1 <- jags(data = jags_data, 
                 model.file = "./R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel1.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b2", "b3", "b4","b5","b6","b7","b8"),
                 n.iter = ni, 
                 n.burnin = nb, 
                 n.chains = nc,
                 n.thin = nt,
                 parallel = TRUE,
                 inits = init_func)
occ_m1 <- sim_fit1
#saveRDS(sim_fit1, file = "./R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Models_Ran_Outputs/Simulations/Test_JAGS_RecoverParams_GlobalDetMod2.Rdata")
```

Evaluate convergence
```{R}
# Evaluating convergence
print(occ_m1)
tracedens_jags(occ_m1, parmfrow = c(3,3))
```

The convergence doesn't look great. THe effective sample size for a0 is too low, and the effective sample size for b2 is just barely above the acceptible threshold. The traceplots for these don't look great. I suspect the a0 didn't converge because we have too many parameters in our model, and b2 didn't converge because I am treating a factor as a numeric value.
DIC 305

**How do you specify a factor variable in a JAGS model?**
**What are the rules of thumb about how many parameters to have in a model??**
**What if I run this with far fewer variables?**

```{R Create Violin Plots Quad Mod}
# Plot posterior distribution
chains_m1 <- jags_df(occ_m1)
# Select the chains for our covariates of interest
chains_viol1 <- chains_m1 %>% select(b1,b2,b3,b4,b5,b6,b7,b8)
# Rename them to be more interpretable
colnames(chains_viol1) <- c("tree richness",
                            "shrub community",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover")
# Pivot this longer so that we can visualize it
chains_viol_long1 <- chains_viol1 %>% pivot_longer(cols = c("tree richness",
                            "shrub community",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover"),names_to = "Parameter", values_to = "Values")

true_vals <- data.frame(beta,covariate=c("tree richness",
                            "shrub community",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover"))
# Create the violin plot
ggplot() +
  geom_violin(data=chains_viol_long1, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Simulated Occupancy Covariates") +
  scale_fill_manual(values = c("tree richness"=palette_8[1], "shrub community"=palette_8[2], "subcanopy height" = palette_8[3], "subcanopy cover" = palette_8[4], "subcanopy complexity" = palette_8[5], "veg density" = palette_8[6], "canopy height" = palette_8[7], "canopy percent cover" = palette_8[8])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_point(data = true_vals, aes(x=covariate, y = beta))+
  theme_minimal()
```

It doesn't look like the model is recovering the parameters well. What does this mean? Why is it not recovering well and why is a0 not converging?


## Try 2: Removing parameters from Model

I am removing b2 (shrub community) because I am not sure how to parameterize it and b6 because I suspect it will be colinear with other variables. Let's see if the model performs better with less complexity. 

```{R Set Up Model,eval = FALSE}
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_unit){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + (b1*tree_rich[i]) + (b3*subcan_h[i]) + (b4*subcan_pc[i]) + (b5*subcan_cx[i]) + (b7*can_h[i]) + (b8*can_pc[i])
      
      cov2[i] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
      
      for(j in 1:n_rep){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0
        
      }
    }
    
    # Priors
    a0 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dnorm(0, 1)
    b3 ~ dnorm(0, 1)
    b4 ~ dnorm(0, 1)
    b5 ~ dnorm(0, 1)
    b7 ~ dnorm(0, 1)
    b8 ~ dnorm(0, 1)
  }
  ", file = "./R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel1.txt")
```

*Next up:* Figure out what the parameterization of these models means
```{r}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z,
  # Detection/nondetection data
  det_data = y,
  # Tree richness
  tree_rich = tree_rich,
  # Subcanopy height
  subcan_h = subcan_h,
  # Subcanopy cover
  subcan_pc = subcan_pc,
  # Subcanopy complexity 
  subcan_cx = subcan_cx,
  # Canoy height
  can_h = can_h,
  # Canopy cover
  can_pc = can_pc,
  n_unit = nP,
  n_rep = nV
)

# Set initial values (optional)
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1),
    b5 = rnorm(1, 0, 1),
    b7 = rnorm(1, 0, 1),
    b8 = rnorm(1, 0, 1)
  )
}

ni <- 10000
nb <- 5000
nc <- 4
nt <- 2

# Run JAGS model on the same model we specify in the Fit_JAGS_Detection_Model2_HabChap markdown
occ_m2 <- jags(data = jags_data, 
                 model.file = "./R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel1.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b3", "b4","b5","b7","b8"),
                 n.iter = ni, 
                 n.burnin = nb, 
                 n.chains = nc,
                 n.thin = nt,
                 parallel = TRUE,
                 inits = init_func)
#saveRDS(sim_fit1, file = "./R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Models_Ran_Outputs/Simulations/Test_JAGS_RecoverParams_GlobalDetMod2.Rdata")
```

Evaluate convergence
```{R}
# Evaluating convergence
print(occ_m2)
tracedens_jags(occ_m2, parmfrow = c(3,3))
```

The model looks better now, the effective sample sizes are all wthin the comfortable range. THe traceplors are still a little bit shakey though. DIC 341