---
title: "Recover Parameters for Simulated Data Occupancy Model"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(coda)
library(janitor)
library(loo)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/6_Function_Scripts/Create_SiteColumn_fromPointID.R")
```


# Simulate Data


```{R generate data, out.width = "50%", out.height = "50%"}
# b1/cov1: tree richness
# b2/cov2: shrub community
# b3/cov3: subcanopy avg height
# b4/cov4: subcanopy percent cover
# b5/cov5: subcanopy veg complexity 
# b6/cov6: veg density 
# b7/cov7: canopy height
# b8/cov8: canopy percent cover

nP <- 107  # number of points
nV <- 6    # number of repeated visits
nC <- 8    # number of covariates
#survey_covs <- 3

# create latent state (z) and detection (y) data storage
z <- NULL        # point-level latent state
y <- array(NA, dim = c(nP, nV))   # point-level observation data

# create separate arrays for each covariate and fill with randomly generated data
# Covariate 1: Tree Richness
set.seed(22)
rich_vals <- c(-0.617, 0.71, 2.03, 3.37) # Create possible values for species richness
tree_rich <- sample(rich_vals, nP, replace =TRUE)
hist(tree_rich, col = cuckoo_palette[1], main = "Distribution of Simulated Native Broadleaf Richness")
# Expected response: linear and positive - higher levels of richness, higher insect abundance - more likely cuckoo presence 

# Covariate 2: Shrub Community
set.seed(3)
com_fs <- c(1:5)
shrub_com <- sample(com_fs, nP, replace = TRUE)
hist(shrub_com, col = cuckoo_palette[1], main = "Distribution of Simulated Shrub Community")
#shrub_com <- as.factor(shrub_com)
# Expected response: higher probability of occ with floodplain and broadleaf communities 

# Covariate 3: Subcanopy Veg Avg Height
set.seed(10)
subcan_h <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_h, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Height")
# Expected response: linear and positive - higher subcanopy veg means more subcanpy habitat and greater prob of occ

# Covariate 4: Subcanopy Veg Percent Cover
set.seed(11)
subcan_pc <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_pc, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Height")
# Expected response: linear and positive - higher subcanopy veg cover means more places to forage and hide and greater prob of occ

# Covariate 5: Subcanopy Veg Complexity 
set.seed(12)
subcan_cx <- rnorm(nP, mean = 1, sd = 1)
hist(subcan_cx, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Subcanopy Complexity")
# Expected response: linear and positive - higher complexity means more places to forage and hide, greater prob of use 

# Covariate 6: Overall Veg Density
set.seed(13)
veg_d <- rnorm(nP, mean = 1, sd = 1)
hist(veg_d, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Total Vegetation Density")
# Expected response: linear and positive - higher veg cover means more places to forage and hide and greater prob of occ

# Covariate 7: Canopy Height
set.seed(14)
can_h <- rnorm(nP, mean = 1, sd = 1)
hist(can_h, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Canopy Height")
# Expected response: selecting for intermediate? want shrubby areas with some trees but not super matrure cottonwood gallery forest ???????????????????????????????????

# Covariate 8: Canopy Percent Cover
set.seed(15)
can_pc <- rnorm(nP, mean = 1, sd = 1)
hist(can_pc, col = cuckoo_palette[1], main = "Distribution of Simulated Scaled Canopy Cover")
# Expected response: also selecting for intermediate? want intermediate levels of canopy cover. since they prefer shrubby habitat. how to parameterize this ??????????????????????????????????? 
# BUT selecting for high canopy cover at a larger scale 

# covariate effects (pull randomly for each covariate)
set.seed(5)
beta <- rnorm(nC, 0, 1)

beta_shrubcom <- rnorm(length(com_fs), 0, 1)

# Baseline probability of presence
# psi0 <- qlogis(0.2) #Original
psi0 <- c()
for (i in 1:length(com_fs)){
  psi0[i] <- (0.2 - beta_shrubcom[i])
}
# Baseline probability of detection
p0 <- qlogis(0.6)

psi <- NULL
for (i in 1:nP){
  psi[i] = plogis(psi0[shrub_com[i]] + (beta[1]*tree_rich[i]) + (beta[3]*subcan_h[i]) + (beta[4]*subcan_pc[i]) + (beta[5]*subcan_cx[i]) + (beta[6]*veg_d[i]) + (beta[7]*can_h[i]) + (beta[8]*can_pc[i]))
}
# Assume equal probability of presence at all sites with this simulation

# get your true presence/non presence values from psi
set.seed(8) # I changed this around until I got a value of presence/absence that looked good to me
z <- rbinom(nP, 1, psi)

# total number of detections across number of visits (nV)
y <- matrix(NA, nP, nV)
# fill in the detection matrix dependent on whether the animal was there (z) and whether it was detected (p0)
# add covariate effects here
set.seed(7)
for (i in 1:nP){
  for (j in 1:nV){
    y[i,j] <- rbinom(1, z[i], plogis(p0))
    # Assume equal probability of detection at all sites and all surveys
  }
}

# b1/cov1: tree richness
# b2/cov2: shrub community
# b3/cov3: subcanopy avg height
# b4/cov4: subcanopy percent cover
# b5/cov5: subcanopy veg complexity 
# b6/cov6: veg density 
# b7/cov7: canopy height
# b8/cov8: canopy percent cover
```

```{R Set Up Model,eval = FALSE}
cat("
  model{
      # Priors
    for (k in 1:com_fs){
      a0[k] ~ dlogis(0, 1)
    }
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1) # took out b2 (shrub communities)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
    b5 ~ dlogis(0, 1)
    b6 ~ dlogis(0, 1)
    b7 ~ dlogis(0, 1)
    b8 ~ dlogis(0, 1)
    
    
    # Likelihood
    # Loop through each sampling unit
    for(i in 1:n_unit){ 
      Z[i] ~ dbern(psi[i])  # Presence or absence at site drawn from a bernoulli distribution with occupancy probability psi dependent on habitat type
      logit(psi[i]) <- a0[shrub_com[i]] + (b1*tree_rich[i]) + (b3*subcan_h[i]) + (b4*subcan_pc[i]) + (b5*subcan_cx[i]) + (b6*veg_d[i]) + (b7*can_h[i]) + (b8*can_pc[i])
      
      cov2[i] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
      
      for(j in 1:n_rep){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0
        
      }
    }
  
    # Derived parameters
    for(i in 1:com_fs){
        occ[i] <- 1/(1 + exp(a0[i])) #Back transform the effect of each intercept 
    }
    # Use a double for loop to derive detection probabilities
    det <- ilogit(b0) # Back transform the probability of detection 
  }
  ", file = "./R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel2_ShrubCommIntercept.txt")
```

```{r}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z,
  # Detection/nondetection data
  det_data = y,
  # Tree richness
  tree_rich = tree_rich,
  # Shrub community
  shrub_com = shrub_com,
  # Number of shrub communities
  com_fs = max(com_fs),
  # Subcanopy height
  subcan_h = subcan_h,
  # Subcanopy cover
  subcan_pc = subcan_pc,
  # Subcanopy complexity 
  subcan_cx = subcan_cx,
  # Overall veg density
  veg_d = veg_d,
  # Canoy height
  can_h = can_h,
  # Canopy cover
  can_pc = can_pc,
  n_unit = nP,
  n_rep = nV
)

# Set initial values (optional)
init_func <- function(){
  list(
    a0 = rnorm(5, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1),
    b5 = rnorm(1, 0, 1),
    b6 = rnorm(1, 0, 1),
    b7 = rnorm(1, 0, 1),
    b8 = rnorm(1, 0, 1)
  )
}

ni <- 10000
nb <- 5000
nc <- 4
nt <- 2

# Run JAGS model on the same model we specify in the Fit_JAGS_Detection_Model2_HabChap markdown
occ_m1 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Occupancy_Submodel/Model_Structure_Files/Test_JAGS_Occ_OccupancyModel2_ShrubCommIntercept.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b3", "b4","b5","b6","b7","b8", "occ", "det"),
                 n.iter = ni, 
                 n.burnin = nb, 
                 n.chains = nc,
                 n.thin = nt,
                 parallel = TRUE,
                 inits = init_func)
#saveRDS(sim_fit1, file = "./R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Models_Ran_Outputs/Simulations/Test_JAGS_RecoverParams_GlobalDetMod2.Rdata")
```

Evaluate convergence
```{R Evaluate Convergence M1}
# Evaluating convergence
print(occ_m1)
# occ now gives us the derived parameters for occupancy at each shrub community (directly interpretable as percent)
tracedens_jags(occ_m1, parmfrow = c(3,3))
```

Evaluate parameter recovery

```{R Create Violin Plots Mod1}
# Plot posterior distribution
chains_m1 <- jags_df(occ_m1)
# Select the chains for our covariates of interest
chains_viol1 <- chains_m1 %>% select(b1,b3,b4,b5,b6,b7,b8)
# Rename them to be more interpretable
colnames(chains_viol1) <- c("tree richness",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover")
# Pivot this longer so that we can visualize it
chains_viol_long1 <- chains_viol1 %>% pivot_longer(cols = c("tree richness",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover"),names_to = "Parameter", values_to = "Values")

true_vals <- data.frame(beta[-2],covariate=c("tree richness",
                            "subcanopy height",
                            "subcanopy cover",
                            "subcanopy complexity",
                            "veg density",
                            "canopy height",
                            "canopy percent cover"))
colnames(true_vals) <- c("beta","covariate")
# Create the violin plot for covariate effects (not including shrub community)
ggplot() +
  geom_violin(data=chains_viol_long1, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Simulated Occupancy Covariates") +
  scale_fill_manual(values = c("tree richness"=palette_8[1], "subcanopy height" = palette_8[3], "subcanopy cover" = palette_8[4], "subcanopy complexity" = palette_8[5], "veg density" = palette_8[6], "canopy height" = palette_8[7], "canopy percent cover" = palette_8[8])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_point(data = true_vals, aes(x=covariate, y = beta))+
  theme_minimal()
```

It doesn't look like the model is recovering the parameters well. What does this mean?

Could it be not recovering the covariates with a weak effect?
* if so, why was subcanopy complexity (with a true strong positive effect) recovered worse than veg density, canopy percent cover, and canopy height (much weaker negative effect)?


```{R Evaluate Recovery of Shrub Intercepts}
# Select the chains for our covariates of interest
chains_viol2 <- chains_m1 %>% select("a0[1]","a0[2]","a0[3]","a0[4]","a0[5]")
colnames(chains_viol2) <- c("shrub_c1","shrub_c2","shrub_c3","shrub_c4","shrub_c5")
# Pivot this longer so that we can visualize it
chains_viol_long2 <- chains_viol2 %>% pivot_longer(cols = c("shrub_c1","shrub_c2","shrub_c3","shrub_c4","shrub_c5"),
                                                   names_to = "Parameter", 
                                                   values_to = "Values")
# Pull out the true values of the beta
true_vals2 <- data.frame(psi0, covariate = c("shrub_c1","shrub_c2","shrub_c3","shrub_c4","shrub_c5"))

ggplot() +
  geom_violin(data=chains_viol_long2, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Shrub Community Baseline Occupancy") +
  scale_fill_manual(values = c("shrub_c1"=palette_5[1], "shrub_c2" = palette_5[2], "shrub_c3" = palette_5[3], "shrub_c4" = palette_5[4], "shrub_c5" = palette_5[5])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_point(data = true_vals2, aes(x=covariate, y = psi0))+
  theme_minimal()
```

THis looks okay, still not super centered around the values I want it to be centered around.Why is it recovering poorly the covariates with a strong effect but recovering those with a weak effect better?

Could it be that I have too many variables in this model?

TODO: 
Google around to see if there's tips on what to do when your model doesn't converge



**What are the rules of thumb about how many parameters to have in a model??**
**What if I run this with far fewer variables?**
- what about a model with four variables? some kind of PCA of canopy composition, a PCA of subcanopy composition, shrub community, and tree richness



