---
title: "Detection Submodel 4"
author: "Anna Kurtin"
date: " 6/28/2024"
output: html_document
---

Pending:

* How does JAGS treat NAs? Ask Colton about this

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(rjags)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(coda)
library(janitor)
library(loo)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
```

Load in formatted data

```{r Load in Data}
full_dat <- readRDS(file ="C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Habitat_Model_Covariates/Full_DetectionDataSCALED_JAGSModels_HabChap.Rdata")
```

Set up data for model 

```{R Specify Data for Model, eval = FALSE}
# Pull out the detections so that you have n_unit and n_site
detections_only <- full_dat[,2:7]

# Have to specify the z data for a true presence if there was a 1 at any time 
z_dat <- rowSums(detections_only, na.rm = T)
z_dat[z_dat > 1] <- 1
z_dat[z_dat == 0] <- NA

# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z_dat,
  # Detection/nondetection data
  det_data = detections_only,
  # Date
  cov1 = full_dat[,21:26],
  # Veg Density - site level
  cov2 = full_dat[,14],
  # Background noise
  cov3 = full_dat[,8:13],
  # Effort 
  cov4 = full_dat[,15:20],
  n_unit = nrow(detections_only),
  n_rep = ncol(detections_only)
)

# Set initial values (optional)
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b2 = rnorm(1, 0, 1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1)
  )
}
```

## Non Quadratic Model

**AKA Model 2**

```{R Set Up Model 2, eval = FALSE}
# b1/cov1: date
# b2/cov2: veg density
# b3/cov3: background noise
# b4/cov4: effort

# Create the jags_model
cat("
  model{
      # Priors
    a0 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1)
    b2 ~ dlogis(0, 1)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
  
    # Loop through each sampling unit
    for(i in 1:n_unit){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0
      
      cov2[i] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
      
      for(j in 1:n_rep){
        cov1[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        cov3[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        cov4[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0 + b1*cov1[i,j] + b2*cov2[i] + b3*cov3[i,j] + b4*cov4[i,j]
        
      }
    }
  }
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Model_Structure_Files/JAGS_HabDetMod2_LogisPriors.txt")
```
```{R Run Model 2}
habdet_m2 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Model_Structure_Files/JAGS_HabDetMod2_LogisPriors.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b2", "b3", "b4"),
                 n.iter = 10000, 
                 n.burnin = 4000, 
                 n.chains = 3,
                 n.thin = 2,
                 #parallel = TRUE,
                 inits = init_func)
```

```{R Evaluate Model 2}
# Evaluating convergence
print(habdet_m2)
tracedens_jags(habdet_m2, parmfrow = c(3,3))
```

```{R Create Violin Plots M2}
# Plot posterior distribution
chains_m2 <- jags_df(habdet_m2)
# Select the chains for our covariates of interest
chains_viol <- chains_m2 %>% select(b1,b2,b3,b4)
# Rename them to be more interpretable
colnames(chains_viol) <- c("date","veg_density","background_noise","effort")
# Pivot this longer so that we can visualize it
chains_viol_long <- chains_viol %>% pivot_longer(cols = c("date","veg_density","background_noise","effort"),names_to = "Parameter", values_to = "Values")

# Create the violin plot
ggplot() +
  geom_violin(data=chains_viol_long, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Detection Covariates") +
  scale_fill_manual(values = c("date"=palette_5[1], "background_noise"=palette_5[2], "effort" = palette_5[3], "veg_density" = palette_5[4])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()
```


```{R}
# Extract samples from the posterior
habdet_m2_s <- jags.samples(habdet_m2$model,
                           c("WAIC","deviance"),
                           type = "mean",
                           n.iter = 5000,
                           n.burnin = 1000,
                           n.thin = 2)
# Duplicate WAIC samples - the WAIC is the posterior mean for that sample?
habdet_m2_s$p_waic <- habdet_m2_s$WAIC
# Calculate a new waic by adding the waic plus the deviance
habdet_m2_s$waic_calc <- habdet_m2_s$deviance + habdet_m2_s$p_waic
# Create a new vector of the sum across each array (sum WAIC, sum deviance, sum p_waic, and sum waic_calc)
tmp <- sapply(habdet_m2_s, sum)
# Pull out the sum of the WAIC we calculated and the p_waic we duplicated
waic_m2 <- round(c(waic = tmp[["waic_calc"]], p_waic = tmp[["p_waic"]]),1)
# Calculate the standard error for WAIC
m2_waic_se <- sd(habdet_m2_s$waic_calc)/(sqrt(length(habdet_m2_s$waic_calc)))

waic_m2
m2_waic_se
# WAIC is 5625.9
# SE is 0.036
```


## Quadratic Model 

**AKA Model 4**

This time properly specified

```{R Set Initial Values M4}
# Set initial values (optional)
init_func4 <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b1Q = rnorm(1, 0 ,1),
    b2 = rnorm(1, 0, 1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1)
  )
}
```



```{R Set Up Model 4, eval = FALSE}
# b1/cov1: date
# b2/cov2: veg density
# b3/cov3: background noise
# b4/cov4: effort

# Create the jags_model
cat("
  model{
      # Priors
    a0 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1)
    b1Q ~ dlogis(0, 1)
    b2 ~ dlogis(0, 1)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
    
    # Loop through each sampling unit
    for(i in 1:n_unit){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0
      
      cov2[i] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
      
      for(j in 1:n_rep){
        cov1[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        cov3[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        cov4[i,j] ~ dnorm(0, 1) # Assuming normal distribution for missing covariates
        
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0 + b1*cov1[i,j] + b1Q*(cov1[i,j]^2) + b2*cov2[i] + b3*cov3[i,j] + b4*cov4[i,j]
        
      }
    }
  } 
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Model_Structure_Files/JAGS_HabDetMod4_QuadDate.txt")
```


```{R}
habdet_m4 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Detection_Submodel/Model_Structure_Files/JAGS_HabDetMod3_QuadDate.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b1Q", "b2", "b3", "b4"),
                 n.iter = 10000, 
                 n.burnin = 4000, 
                 n.chains = 3,
                 n.thin = 2,
                 #parallel = TRUE,
                 inits = init_func4)
# Unused initial value for b1Q???
print(habdet_m4)
```

Why is this model giving me an error that there is an unused initial value bor B1Q? How can it converge without an initial value (probably just radomly chooses one from a normal (?) distribution)