---
title: "Basic Occupancy Model in JAGS"
author: "Anna Kurtin"
date: "3/31/2024"
output: html_document
---

Combine with your project proposal to make the intro for this first draft? 

JAGS is a more flexible way to run bayesian models. When you build JAGS code, you write out your full model and specify it's formulas, as opposed to more "black box" packages such as unmarked and ubms. 
[explanation of jags vs stan](https://bookdown.org/steve_midway/BHME/Ch4.html)

[demo of ubms vs jags](https://github.com/kenkellner/ubms/blob/master/vignettes/JAGS-comparison.Rmd)

JAGS doesn't, however, allow you to build your own MCMC sampler. **Which one does it use?** In order to do so, you will need to use a software such as NIMBLE. JAGS also can't handle NA data (?).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(unmarked)
library(jagsUI)
library(tidyverse)
```

# Basic Occupancy Model 1: Known Parameters
For the first simulation, we'll run a jags model assuming that we know the parameters behind the model. Then, we'll run a jags model where we don't know the parameters, only the data.

First, we are going to simulate the data and model assuming we know the parameter
```{r Simulate Data 1}
set.seed(123)

n_unit <- 85
n_rep <- 3

# Simulate covariates
# Site level covariate: probability of cottonwood (Populus sp.) presence
cov1 <- round(rbeta(n_unit, .70, .15) ,2)
# Site level covariate: percent of lidar returns at the median height
cov2 <- round(rbeta(n_unit, .50, .20),2)
# Survey level covariate: dB of background noise
cov3 <- matrix(runif(n_unit * n_rep, min = -70, max = -17), nrow = n_unit) # Visibility
# Center and scale this to perform better in our models
cov3_fin <- scale(cov3, center = TRUE)
cov3_alt <- matrix(rnorm(n_unit * n_rep, mean = 1, sd = .5), nrow = n_unit) # Visibility

# Define true parameter values for simulating data
a0_true <- .20 # baseline probability of occupancy (originally -1)
a1_true <- 0.3 # effect of cottonwood presence on occupancy 
a2_true <- 0.5 # effect of median height veg density on occupancy 
b0_true <- .60 # baseline probability of detection (originally -2)
b1_true <- 0.8 # not used
b2_true <- 0.5 # effect of background noise on detection

# Simulate occupancy probability (psi) at each unit
psi <- plogis(a0_true + a1_true * cov1 + a2_true * cov2)
# When we set the values of cov1 and cov2 with raw counts or percentages, these are all 1. This is why we scale things?

# Simulate true presence/absence
Z <- rbinom(n_unit, 1, psi)

# Initialize arrays to store detection probabilities and detection data
theta <- array(NA, dim = c(n_unit, n_rep))
det_data <- array(NA, dim = c(n_unit, n_rep))

# Simulate detection probability (theta) and detection data (det_data)
## Have the same covariate in the detection and the biological model?
for (i in 1:n_unit) {
  theta[i, ] <- plogis(b0_true + b2_true * cov3_alt[i, ])
  det_data[i, ] <- rbinom(n_rep, 1, Z[i] * theta[i, ])
}
# Detection probability is generally pretty low (around 18%) so I changed cov3 to give a higher detection probability 
mean(theta[,1]) #18% with cov3_alt. 12 with cov3_fin - this is zero-inflated with cov3_fin
# How to represent dB without zero-inflating data?
# Come back to adjust cov3

```

```{r Run JAGS Model 1}
# Create the jags_model
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_unit){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + a1*cov1[i] + a2*cov2[i]
      for(j in 1:n_rep){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0 + b2*cov3[i,j]
      }
    }
    # Priors
    a0 ~ dnorm(0, 0.01)
    a1 ~ dnorm(0, 0.01)
    a2 ~ dnorm(0, 0.01)
    b0 ~ dnorm(0, 0.01)
    b2 ~ dnorm(0, 0.01)
  }
  ", file = "Simulation_JAGS_Occ_Model1.txt")

# Create a list of data for JAGS
jags_data1 <- list(
  Z = Z,
  det_data = det_data,
  cov1 = cov1,
  cov2 = cov2,
  cov3 = cov3_alt,
  n_unit = n_unit,
  n_rep = n_rep
)

# Set initial values (optional)
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    a1 = rnorm(1, 0, 1),
    a2 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b2 = rnorm(1, 0, 1)
  )
}

# Run JAGS model
jags_fit1 <- jags(data = jags_data1, 
                 model.file = "Simulation_JAGS_Occ_Model1.txt",
                 parameters.to.save = c("a0", "a1", "a2", "b0", "b2"),
                 n.iter = 10000, 
                 n.burnin = 2000, 
                 n.chains = 3,
                 inits = init_func)

jags_fit1
```


```{r look at JAGS Model 1 Results}
# Get a summary of the results
summary(jags_fit1)

# Plot the results
plot(jags_fit1)
# walk through what each of these look like?
# The plots look grassy - good. What does Rhat mean? 
# Are the density plots the posterior distribution?
# Extract MCMC chains
#chains <- as.mcmc(jags_fit1)


```

Looking at model 1 trace plots and density plots:

* a0 - the baseline probability of occupancy is estimated to be 0.57, with a credible interval from -1.23 to 2.5.

* a1 - the effect of probability of populus presence on occupancy. Mean value is -0.046, with a credible interval from -1.96 to 1.71

* a2 - the effect of vegetation density on occupancy. Mean value is 0.79, with a credible interval from -0.65 to 2.19.

* b0 - the baseline probability of detection is estimated to be 0.51, with a credible interval from -0.15 to 1.18.

* b2 - the effect of sonic masking on detection probability. Mean value is 0.40, with a credible interval -0.25 to 1.06.

**Does is make sense for these to be more and less than 1? Am I interpreting these right?**


# Basic Occupancy Model 2: Unknown Parameters

Now, we are going to run the JAGS model assuming unknown parameters. 

```{r Simulate Data for JAGS}
# How to format data for JAGS? can it only be from 0-1

# JAGS only takes numeric data (int or factor)
set.seed(123)
# generate a list of point IDs
#point_id <- c(1:85)
#n_site <- length(point_id)
n_site <- 85
n_visit <- 3

# Simulate covariates
# Site level covariate: probability of cottonwood (Populus sp.) presence
populus1 <- round(rbeta(n_site, .70, .15) ,2)
# Discrete: simulate whether cottonwood was present or not at the sites (estimate a 30% probability that cottonwood present) - this is more like what our data actually looks like
populus2 <- rbinom(n = n_site, size = 1, prob = .35)
# Try scaling this for our model?
populus3 <- scale(populus2,center = TRUE)
# TRY USING POPULUS2 FIRST

# Site level covariate: subcanopy density - percent of lidar returns at the median height
veg_dense1 <- round(rbeta(n_site, .50, .20),2)
# Continuous: Simulate GEDI percent of returns coming from 50% height
veg_dense2 <- round(runif(n = n_site, min = 10, max = 80),2)
veg_dense3 <- scale(veg_dense2, center = TRUE)

# Survey level covariate: dB of background noise
background_db1 <- matrix(rnorm(n_site * n_visit, mean = 1, sd = .5), nrow = n_site)
background_db2 <- matrix(runif(n_site * n_visit, min = -70, max = -17), nrow = n_site)
# Center and scale this to perform better in our models
background_db3 <- scale(background_db2, center = TRUE)
 # Visibility

# Simulate 3 sampling periods of detection data based on prelim detections from habitat points
# 85 total habitat points, 18 with detections, 21% naive occupancy 
detected_visit1 <-rbinom(n = n_site,size = 1, prob = .21)
detected_visit2 <-rbinom(n = n_site,size = 1, prob = .21)
detected_visit3 <-rbinom(n = n_site,size = 1, prob = .21)
detection_hist <- cbind(detected_visit1,detected_visit2,detected_visit3)
```


```{r Run JAGS Model Unknown Params, eval = FALSE}
# Create the jags_model
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_site){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli dist with occupancy probability psi  
      logit(psi[i]) <- a0 + a_pop*populus[i] + a_vd*veg_dense[i]
      
      # Loop through each visit
      for(j in 1:n_visit){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data 
        logit(theta[i,j]) <- b0 + b_masking*background_db[i,j]
      }
    }
    # Priors
    a0 ~ dnorm(0, 0.01)
    a_pop ~ dnorm(0, 0.01)
    a_vd ~ dnorm(0, 0.01)
    b0 ~ dnorm(0, 0.01)
    b_masking ~ dnorm(0, 0.01)

  }", file = ".//Model_Files//Simulation_JAGS_Occ_Model4.txt")

# Create a list of data for JAGS
jags_data <- list(
  det_data = detection_hist,
  populus = populus1,
  veg_dense = veg_dense1,
  background_db = background_db1,
  n_site = n_site,
  n_visit = n_visit
)

# Establish which parameters for the model to monitor
params <- c("Z","a0", "a1", "a2", "b0", "b2") 

# Bayesian models need a place to start for latent variables 
# Set initial values (optional)
zst <- rep(1,n_site)
init_func <- function(){
  list(
    Z = zst,
    a0 = rnorm(1, 0, 1),
    a_pop = rnorm(1, 0, 1),
    a_vd = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b_masking = rnorm(1, 0, 1)
  )
}

# Run JAGS model
jags_fit <- jags(data = jags_data, 
                 model.file = "./Model_Files/Simulation_JAGS_Occ_Model4.txt",
                 parameters.to.save = c("a0", "a_pop", "a_vd", "b0", "b_masking"), # These are the covariates that will appear in your traceplots
                 n.iter = 3000, 
                 n.burnin = 800, 
                 n.chains = 3,
                 inits = init_func)


# Look at how the model performed
jags_fit

# Look at just the values
summary(jags_fit)

# Plot the results
plot(jags_fit)
# this one looks like it failed to converge - why?
```

**Can't get this part to run**
```{r Run JAGS Model 3, eval = FALSE}
# Create the jags_model
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_site){
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli dist with occupancy probability psi  
      logit(psi[i]) <- a0 + a_pop*populus[i] + a_vd*veg_dense[i]
      
      # Loop through each visit
      for(j in 1:n_visit){
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Create detection data 
        logit(theta[i,j]) <- b0 + b_masking*background_db[i,j]
      }
    }
    # Priors
    a0 ~ dnorm(0, 0.01)
    a_pop ~ dnorm(0, 0.01)
    a_vd ~ dnorm(0, 0.01)
    b0 ~ dnorm(0, 0.01)
    b_masking ~ dnorm(0, 0.01)

  }", file = ".//Model_Files//Simulation_JAGS_Occ_Model5.txt")

# Create a list of data for JAGS
# running this on realistic, unscaled data
jags_data2 <- list(
  det_data = detection_hist,
  populus = populus2,
  veg_dense = veg_dense2,
  background_db = background_db3,
  n_site = n_site,
  n_visit = n_visit
)

# Establish which parameters for the model to monitor
params <- c("Z","a0", "a_pop", "a_vd", "b0", "b_masking") 

# Bayesian models need a place to start for latent variables 
# Set initial values (optional)
zst <- rep(1,n_site)
init_func <- function(){
  list(
    Z = zst,
    a0 = rnorm(1, 0, 1),
    a_pop = rnorm(1, 0, 1),
    a_vd = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b_masking = rnorm(1, 0, 1)
  )
}

# Run JAGS model
jags_fit2 <- jags(data = jags_data2, 
                 model.file = "./Model_Files/Simulation_JAGS_Occ_Model5.txt",
                 parameters.to.save = c("a0", "a_pop", "a_vd", "b0", "b_masking"), # These are the covariates that will appear in your traceplots
                 n.iter = 3000, 
                 n.burnin = 800, 
                 n.chains = 3,
                 inits = init_func)


# Look at how the model performed
jags_fit2

# Look at just the values
summary(jags_fit2)

# Plot the results
plot(jags_fit2)
# this one looks like it failed to converge - why?
```
This model worked for the standard variables. What if we use them on more realistic variables that more closely reflect the data?


# Code Graveyard
```{r CODE GRAVEYARD, inclue = FALSE, eval = FALSE}
# # Set initial values (optional)
# init_func <- function(){
#   list(
#     a0 = rnorm(1, 0, 1),
#     a1 = rnorm(1, 0, 1),
#     a2 = rnorm(1, 0, 1),
#     b0 = rnorm(1, 0, 1),
#     b2 = rnorm(1, 0, 1)
#   )
# }

# # Leaving this off for now
# # date of first survey ranging from June 1 (J.D. 152) to Aug 1 (J.D. 213)
# julian_date1 <- round(runif(n=n_site, min = 152, max = 213),0)
# # date of second survey: two days apart
# julian_date2 <- julian_date1 + 2
# # date of last survey: two days apart
# julian_date3 <- julian_date2 + 2
# # Make this into a matrix
# date_matrix <- cbind(julian_date1,julian_date2,julian_date3)


# Combine into a list
# data_m2 <- list(detection_hist = detection_hist, 
#              populus = populus, 
#              percent_veg_50 = percent_veg_50, 
#              background_db = background_db, 
#              date_matrix = date_matrix)
# Initialize true presence/absence with naive occupancy 
z <- rbinom(n_unit, 1, .21)
cov3 <- matrix(rnorm(n_site * n_rep, mean = 1, sd = .5), nrow = n_site) # Visibility

```

Example of how to parameterize
```{r JAGS/BUGS Model, eval = FALSE}
# Parameterization of a single-species, single season occupancy model using JAGS/BUGS

# Loop through each study site
for(i in 1:n_site){
  # Calculate probability of the presence or absence of the species at this site
  Z[i] ~ dbern(psi[i]) 
  # Set up the logistic regression model for the probability of occupancy in the sampling unit i
  logit(psi[i]) <- a0 + a1*cov1[i] + a2*cov2[i] 
  
  # Loop through each replicate at the site
  for(j in 1:n_rep){
    # Calculate the probability of detection at this visit - estimated from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta. If Z[i] is 0, detection will also be 0
    det_data[i,j] ~ dbern(Z[i]*theta[i,j])  
    # Set up the logistic regression model for detection probability 
    logit(theta[i,j]) <- b0 + b1*cov1[i] + b2*cov3[i,j]
  }
}
```


n_unit: number of study sites/units 
i: the index for the study site
Z: the presence or absence of a species at the site 
psi: a vector with the value for psi (probability of occupancy)
theta: probability of detection 
a0, a1, a2: coefficients, we want to estimate it's effect on presence/absence (a0 is the intercept and a1 and a2 are the effect of the first and second covariates)
cov1[i]: a site-level covariate (like habitat variable) for sampling unit i
cov2[i]: a site-level covariate (like habitat variable) for sampling unit i
n_rep: number of replicates at each site/unit
j: the index for the replicate
det_data[i,j]: the detection/nondetection of the species in replicate j within sampling unit i
b0. b1, b2: coefficients, we want to estimate it's effect on detection/nondetection
cov1[i]: a survey level covariate for sampling unit i and replicate j
cov2[i]: a survey level covariate for sampling unit i and replicate j


```{r Simulate Data}
set.seed(123)
#### Simulate data ####
colnames <- c("point_id","pres_visit1","pres_visit2","pres_visit3","pres_populus","percent_veg_50","background_db","date1","date2","date3")

# generate a list of point IDs
point_id <- c(1:85)
n_site <- length(point_id)
# Simulate 3 sampling periods of detection data based on prelim detections from habitat points
# 85 total habitat points, 18 with detections, 21% naive occupancy 
detected_visit1 <-rbinom(n = n_site,size = 1, prob = .21)
detected_visit2 <-rbinom(n = n_site,size = 1, prob = .21)
detected_visit3 <-rbinom(n = n_site,size = 1, prob = .21)
detection_hist <- cbind(point_id,detected_visit1,detected_visit2,detected_visit3)
n_rep <- 3

# Site-level covariates
# Discrete: simulate whether cottonwood was present or not at the sites (estimate a 30% probability that cottonwood present)
populus <- rbinom(n = n_site, size = 1, prob = .35)
# Continuous: Simulate GEDI percent of returns coming from 50% height
percent_veg_50 <- round(runif(n = n_site, min = 10, max = 80),2)

# Survey-level covariates
# Sonic masking could go either way depending on how I want to group it but we're just doing survey level for now
background_db <- matrix(round(runif(n = n_site * n_rep, min = -70, max = -17),2), nrow = n_site)

# date of first survey ranging from June 1 (J.D. 152) to Aug 1 (J.D. 213)
julian_date1 <- round(runif(n=n_site, min = 152, max = 213),0)
# date of second survey: two days apart
julian_date2 <- julian_date1 + 2
# date of last survey: two days apart
julian_date3 <- julian_date2 + 2
# Make this into a matrix
date_matrix <- cbind(julian_date1,julian_date2,julian_date3)


# Combine into dataframe
# Combine lists into a data frame
df <- data.frame(
  point_id = point_id,
  pres_visit1 = pres_visit1,
  pres_visit2 = pres_visit2,
  pres_visit3 = pres_visit3,
  pres_populus = populus,
  percent_veg_50 = percent_veg_50,
  background_db = background_db,
  date1 = julian_date1,
  date2 = julian_date2,
  date3 = julian_date3
)
```

```{r Simulate Data 2}
# JAGS only takes numeric data (int or factor)

# generate a list of point IDs
point_id <- c(1:85)
n_site <- length(point_id)
n_visit <- 3
# Simulate 3 sampling periods of detection data based on prelim detections from habitat points
# 85 total habitat points, 18 with detections, 21% naive occupancy 
detected_visit1 <-rbinom(n = n_site, size = 1, prob = .21)
detected_visit2 <-rbinom(n = n_site, size = 1, prob = .21)
detected_visit3 <-rbinom(n = n_site, size = 1, prob = .21)
detection_hist <- cbind(detected_visit1,detected_visit2,detected_visit3)


# Site-level covariates
# Discrete: simulate whether cottonwood was present or not at the sites (estimate a 30% probability that cottonwood present)
populus <- rbinom(n = n_site, size = 1, prob = .35)
#populus <- scale(populus)
# Continuous: Simulate GEDI percent of returns coming from 50% height
percent_veg_50 <- round(runif(n = n_site, min = 10, max = 80),2)
#percent_veg_50 <- scale(percent_veg_50)

# Survey-level covariates
# Sonic masking could go either way depending on how I want to group it but we're just doing survey level for now
background_db <- matrix(round(runif(n = n_site * n_visit, min = -70, max = -17),2), nrow = n_site)
#background_db <- scale(background_db)

```

Implementing model: this was my failed attempt to create a JAGS model 
```{r Run JAGS Model 2}
# Old code that used to work
# Create the jags_model
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_site){
      z[i] ~ dbern(psi[i])  # State model (occupancy)
      logit(psi[i]) <- a0 + a1*populus[i] + a2*percent_veg_50[i]
      
      # Loop through each visit
      for(j in 1:n_visit){
          C[i,j] ~ dbin(p[i,j], z[i]) # detection model
        logit(p[i,j]) <- b0 + b2*background_db[i,j]
        
      }
    }
      
    # Priors
    a0 ~ dnorm(0, 0.01)
    a1 ~ dnorm(0, 0.01)
    a2 ~ dnorm(0, 0.01)
    b0 ~ dnorm(0, 0.01)
    b2 ~ dnorm(0, 0.01)
  }
  ", file = "Simulation_JAGS_Occ_Model2.txt")

# Create a list of data for JAGS
jags_data2 <- list(
  C = detection_hist,
  populus = populus,
  percent_veg_50 = percent_veg_50,
  background_db = background_db, 
  n_site = n_site,
  n_visit = n_visit
)

# Establish which parameters for the model to monitor
params <- c("a0", "a1", "a2", "b0", "b2") 

# Bayesian models need a place to start for latent variables 
zst <- apply(jags_data2$C, 1, max) #starting values for our latent z variable 
# take the C data, take the max value and start at place 1
inits <- function(){
  list(z = zst, 
       a0 = rnorm(1, 0, 1),
       a1 = rnorm(1, 0, 1),
       a2 = rnorm(1, 0, 1),
       b0 = rnorm(1, 0, 1),
       b2 = rnorm(1, 0, 1))} #starting values

# # Set initial values (optional)
# init_func <- function(){
#   list(
#     a0 = rnorm(1, 0, 1),
#     a1 = rnorm(1, 0, 1),
#     a2 = rnorm(1, 0, 1),
#     b0 = rnorm(1, 0, 1),
#     b2 = rnorm(1, 0, 1)
#   )
# }

# Run JAGS model
jags_fit2 <- jags(data = jags_data2, 
                 model.file = "Simulation_JAGS_Occ_Model2.txt",
                 parameters.to.save = c("a0", "a1", "a2", "b0", "b2"),
                 n.iter = 10000, 
                 n.burnin = 2000, 
                 n.chains = 3,
                 inits = inits)

# Look at how the model performed
jags_fit2

# Look at just the values
summary(jags_fit2)

# Plot the results
plot(jags_fit2)
# this one looks like it failed to converge - why?
```


```{r Run JAGS Model 3}
# Create the jags_model
cat("
  model{
    # Loop through each sampling unit
    for(i in 1:n_site){
      z[i] ~ dbern(psi[i])  # State model (occupancy)
      logit(psi[i]) <- a0 + a1*populus[i] + a2*percent_veg_50[i]
      
      # Loop through each visit
      for(j in 1:n_visit){
        C[i,j] ~ dbern(p[i,j]*z[i]) # detection model
        logit(p[i,j]) <- b0 + b2*background_db[i,j]
      }
    }
      
    # Priors
    a0 ~ dnorm(0, 0.01)
    a1 ~ dnorm(0, 0.01)
    a2 ~ dnorm(0, 0.01)
    b0 ~ dnorm(0, 0.01)
    b2 ~ dnorm(0, 0.01)
  }
  ", file = "Simulation_JAGS_Occ_Model3.txt")

# Create a list of data for JAGS
jags_data2 <- list(
  C = detection_hist,
  populus = populus,
  percent_veg_50 = percent_veg_50,
  background_db = background_db, 
  n_site = n_site,
  n_visit = n_visit
)

# Establish which parameters for the model to monitor
params <- c("z","a0", "a1", "a2", "b0", "b2") 

# Bayesian models need a place to start for latent variables 
#zst <- apply(jags_data2$C, 1, max) #starting values for our latent z variable 
# Bayesian models need a place to start for latent variables 
# zst <- numeric(n_site)  # Create empty vector to store initial values for z
# for (i in 1:n_site) {
#   zst[i] <- max(jags_data2$C[i, ])
# }
zst <- rep(1,n_site)
# take the C data, take the max value and start at place 1
inits_func <- function (){
  list(z = zst, 
       a0 = rnorm(1, 0, 1),
       a1 = rnorm(1, 0, 1),
       a2 = rnorm(1, 0, 1),
       b0 = rnorm(1, 0, 1),
       b2 = rnorm(1, 0, 1))} #starting values


# Run JAGS model
jags_fit2 <- jags(data = jags_data2, 
                 model.file = "Simulation_JAGS_Occ_Model3.txt",
                 parameters.to.save = c("z","a0", "a1", "a2", "b0", "b2"),
                 n.iter = 10000, 
                 n.burnin = 2000, 
                 n.chains = 3,
                 inits = inits_func)

# Look at how the model performed
jags_fit2

# Look at just the values
#summary(jags_fit2)

# Plot the results
#plot(jags_fit2)
# this one looks like it failed to converge - why?
```