---
title: "Full Occupancy Model 3"
author: "Anna Kurtin"
date: "2024-08-08"
output: html_document
---
This is the third iteration of the full occupacy model. From each of the submodels, I carried forward any covariate with f statistic > .70. I assessed covariance of the parameters in the previous model and assessed covariate stability, and discarded subcanopy average height (a5) from this model. This covariate had unstable signs of the parameter effects and was strongly colinear with the parameter estimates of percent subcanopy core (f = -0.483; meaning you can't increase or decrease the effect of one without decreasing or increasing the effect of another).

I also would like to take a closer look at the covariates used in this model and see if I can detect whether one has a "tail" that is causing the intercept for psi/prob of use to be biased lower.

**Model structure:**

State model with six covariates:

* a1: tree species richness on point scale (from submodel A2)

* a2: percent canopy on landscape scale (from submodel B2) 

* a3: percent subcanopy on landscape scale (from submodel B2) 

* a4: percent subcanopy on core scale (from submodel C5)

* a6: residual variation in total veg st dev regressed on canopy height (from submodel C5)

* a7: floodplain dominant shrub community type (from submodel A1)


$$
z_i \sim Bernoulli(\psi | \alpha_{1}, \alpha_{2}, \alpha_{3}, \alpha_{4}, \alpha_{6}, \alpha_{7})
$$ 


Process model with four covariates:  

* b1: date

* b2: quadratic effect of date

* b3: average background decibels

* b4: survey effort

$$
y_{ij} \sim Bernoulli(p| \beta_{1}, \beta_{2}^2,\beta_{3}, \beta_{4})
$$

**Handling of NAs:** omission for detections, imputation for habitat covariates

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(rjags)
library(tidyverse)
library(jagshelper)
library(corrgram)
library(ggplot2)
library(beepr)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
```

```{r Read in Data}
full_dat <- read.csv("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Habitat_Model_Covariates/HabChap_DetOccCovsFull_SCALED_7-30.csv")

detections <- full_dat[,2:7]

# Have to specify the z data for a true presence if there was a 1 at any time 
z_dat <- rowSums(detections, na.rm = T)
z_dat[z_dat > 1] <- 1
z_dat[z_dat == 0] <- NA

# get the length of non-nas in the detection data
get_na <- function(x){(max(which(!is.na(x))))}
# Apply this to our data
miss <- apply(detections, 1, get_na)
```

```{R Visualize Correlations in Data}
corr_dat <- full_dat %>% select(tree_spp_rich, pct_can_landsc, pct_subcan_landsc, pct_subcan_core, veg_sd_resid, floodplain_shrub)
corrgram(corr_dat, order=TRUE, lower.panel=panel.cor,
         upper.panel=panel.pts, text.panel=panel.txt,
         main="Correlations in Habitat Covariates")
```

```{R Look at Data Distributions}
hist(full_dat$tree_spp_rich, main = "Decid tree Spp Richness")
hist(full_dat$pct_can_landsc, breaks = 50, main = "Pct Can Landsc") # has a big positive tail - wouldn't this counter balance the negative tail of the veg sd?
hist(full_dat$pct_subcan_landsc, breaks = 50, main = "Pct Subcan Landsc")
hist(full_dat$pct_subcan_core, breaks = 50, main = "Pct Subcan Core")
hist(full_dat$veg_sd_resid, breaks = 50, main = "Total Veg SD Residuals") # had a big negative tail, could this be what is biasing the mean occupancy low?
hist(full_dat$floodplain_shrub, main = "Floodplain Shrub Community")
```

```{R}
full_dat$bbcu <- apply(full_dat[, 2:7], 1, max, na.rm = TRUE)
# How does points with BBCU vs not correspond to total veg sd resid?
plot(full_dat$bbcu~full_dat$veg_sd_resid)
#fit <- glm(bbcu ~ veg_sd_resid, data = full_dat, family = "binomial")

ggplot(full_dat, aes(x = veg_sd_resid, y = bbcu)) +
  geom_point(alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE, color = "blue") +  # Logistic regression curve
  labs( x = "Vegetation SD Residuals", y = "BBCU Presence (1) / Absence (0)") +
  theme_minimal()

plot(full_dat$bbcu~full_dat$pct_can_landsc)
ggplot(full_dat, aes(x = pct_can_landsc, y = bbcu)) +
  geom_point(alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE, color = "blue") +  # Logistic regression curve
  labs( x = "Percent Canopy Landscape", y = "BBCU Presence (1) / Absence (0)") +
  theme_minimal()
```

```{R Setup JAGS Data}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z_dat,
  # Detection/nondetection data
  det_data = detections,
  # Native broadleaf tree spp richness
  cova1 = full_dat$tree_spp_rich,
  # % canopy landscape
  cova2 = full_dat$pct_can_landsc,
  # % subcanopy landscape
  cova3 = full_dat$pct_subcan_landsc,
  # % subcanopy core
  cova4 = full_dat$pct_subcan_core,
  # residual veg sd
  cova6 = full_dat$veg_sd_resid,
  # floodplain community
  cova7 = full_dat$floodplain_shrub,
  # Date
  covb1 = full_dat[,35:40],
  # Background noise
  covb3 = full_dat[,22:27],
  # Effort 
  covb4 = full_dat[,29:34],
  n_unit = nrow(detections),
  miss = miss
)

```


```{R}
# Set initial values 
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    a1 = rnorm(1, 0, 1),
    a2 = rnorm(1, 0, 1),
    a3 = rnorm(1, 0, 1),
    a4 = rnorm(1, 0, 1),
    a6 = rnorm(1, 0, 1),
    a7 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b1Q = rnorm(1, 0 ,1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1)
  )
}
```

```{R Model 3 Structure, eval = FALSE}
cat("
  model{
        # Priors
    a0 ~ dlogis(0, 1)
    a1 ~ dlogis(0, 1)
    a2 ~ dlogis(0, 1)
    a3 ~ dlogis(0, 1)
    a4 ~ dlogis(0, 1)
    a6 ~ dlogis(0, 1)
    a7 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1)
    b1Q ~ dlogis(0, 1)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
    
    # Likelihood
    # Loop through each sampling unit
    for(i in 1:n_unit){
      # Impute missing habitat data
      cova1[i] ~ dnorm(0, 1)
      cova2[i] ~ dnorm(0, 1)
      cova3[i] ~ dnorm(0, 1)
      cova4[i] ~ dnorm(0, 1)
      cova6[i] ~ dnorm(0, 1)
      cova7[i] ~ dnorm(0, 1)
       
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + a1*cova1[i] + a2*cova2[i] + a3*cova3[i] + a4*cova4[i] + a6*cova6[i] + a7*cova7[i]
      
      for(j in 1:miss[i]){
       # Impute missing detection data
        covb1[i,j] ~ dnorm(0, 1)
        covb3[i,j] ~ dnorm(0, 1)
        covb4[i,j] ~ dnorm(0, 1)
         
        det_data[i,j] ~  dbern(mu_p[i,j])  # Create probability density drawn for each site
        mu_p[i,j] <- Z[i]*theta[i,j]
        logit(theta[i,j]) <- p[i,j]
        p[i,j] <- b0 + b1*covb1[i,j] + b1Q*(covb1[i,j]^2) + b3*covb3[i,j] + b4*covb4[i,j]
        # Create detection data without NAs
        det_na[i,j] <- det_data[i,j]
         
        #Calculate Bayes p-value - plug the parameters into the equation for each sample of the posterior and use this to create new observations
        Presi[i,j] <- (det_na[i,j] - theta[i,j])^2 # Calculate squared residual of observed data
        y_new[i,j] ~ dbern(mu_p[i,j])             # Simulate observed data
        Presi_new[i,j] <- (y_new[i,j] - theta[i,j])^2 # Calculate squared residual error of simulated data
      }
    }
    
    # Derived parameters
    ## Back transform the parameters into probabilities
    pct_can_core <- ilogit(a1)
    pct_subcan_core <- ilogit(a2)
    
    ## Calculate sum of squared residual errors for observed 
    for(i in 1:n_unit){
       sobs[i] <- sum(Presi[i,1:miss[i]])
       ssim[i] <- sum(Presi_new[i,1:miss[i]])
    }
    
    SSEobs <- sum(sobs[])
    SSEsim <- sum(ssim[])
    p_val <- step(SSEsim - SSEobs)  
    
  }
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Full_Occupancy_Model/Model_Structure_Files/JAGS_GlobMod3.txt")

```

```{R Run Model, results = FALSE}
fit_g3 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Full_Occupancy_Model/Model_Structure_Files/JAGS_GlobMod3.txt",
                 parameters.to.save = c("a0", "a1", "a2", "a3", "a4", "a6", "a7", "b0", "b1", "b1Q", "b3", "b4","p_val"),
                 n.iter = 10000, 
                 n.burnin = 4000, 
                 n.chains = 3,
                 n.thin = 2,
                 inits = init_func)
beep(sound = 3)
#Drawing inference from this - need to back transform the parameters
```

```{R}
summary(fit_g3)
tracedens_jags(fit_g3, parmfrow = c(3,3))
```


* cova1: Native broadleaf tree spp richness

* cova2: % canopy landscape

* cova3: % subcanopy landscape

* cova4: % subcanopy core

* cova6: residual veg sd

* cova7: floodplain shrubs dominant community



```{R Create Violin Plots, echo = FALSE}
chains_g3 <- jags_df(fit_g3)
# Select the chains for our covariates of interest
chains_viol <- chains_g3 %>% select(a0, a1,a2, a3, a4, a6, a7)
# Rename them to be more interpretable
colnames(chains_viol) <- c("psi","tree spp richness", "% can landsc", "% subcan landsc", "% subcan core", "residual veg sd", "floodplain shrub")
# Pivot this longer so that we can visualize it
chains_viol_long <- chains_viol %>% pivot_longer(cols = c("psi", "tree spp richness", "% can landsc", "% subcan landsc", "% subcan core", "residual veg sd", "floodplain shrub"),names_to = "Parameter", values_to = "Values")

# Create the violin plot
ggplot() +
  geom_violin(data=chains_viol_long, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Global Model Covariates") +
  scale_fill_manual(values = c("psi" = palette_8[8], "tree spp richness"=palette_8[1], "% can landsc" = palette_8[2], "% subcan landsc" = palette_8[3], "% subcan core" = palette_8[4],  "residual veg sd" = palette_8[6], "floodplain shrub" = palette_8[7])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()

chains_beta <- chains_g3 %>% select(b1, b1Q, b3, b4)
colnames(chains_beta) <- c("date", "quadratic date", "background dB", "effort")
chains_beta_l <- chains_beta %>% pivot_longer(cols = c("date", "quadratic date", "background dB", "effort"),names_to = "Parameter", values_to = "Values")
# Create the violin plot
ggplot() +
  geom_violin(data=chains_beta_l, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Detection Covariates") +
  scale_fill_manual(values = c("date" = palette_5[1], "quadratic date" = palette_5[2], "background dB" = palette_5[3], "effort" = palette_5[4])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()
```


```{R WAIC G3, echo = FALSE}
# Extract samples from the posterior
g3_s <- jags.samples(fit_g3$model,
                           c("WAIC","deviance"),
                           type = "mean",
                           n.iter = 5000,
                           n.burnin = 1000,
                           n.thin = 2)
beep()
# Duplicate WAIC samples - the WAIC is the posterior mean for that sample?
g3_s$p_waic <- g3_s$WAIC
# Calculate a new waic by adding the waic plus the deviance
g3_s$waic_calc <- g3_s$deviance + g3_s$p_waic
# Create a new vector of the sum across each array (sum WAIC, sum deviance, sum p_waic, and sum waic_calc)
tmp <- sapply(g3_s, sum)
# Pull out the sum of the WAIC we calculated and the p_waic we duplicated
waic_g3 <- round(c(waic = tmp[["waic_calc"]], p_waic = tmp[["p_waic"]]),1)
# Calculate the standard error for WAIC
g3_waic_se <- sd(g3_s$waic_calc)/(sqrt(length(g3_s$waic_calc)))
```


```{R Look at Parameter Correlation}
# Create a matrix of the mcmc chains for each parameter
mat <- cbind(fit_g3$sims.list$a0,fit_g3$sims.list$a1,fit_g3$sims.list$a2,
             fit_g3$sims.list$a3,fit_g3$sims.list$a4,
             fit_g3$sims.list$a6,fit_g3$sims.list$a7)
# run a correlation test on the chains - here, I'm looking for values that are larger than other values
r <- cor(mat)
# look at the information contained within the MCMC chains, if the last eigenvector is close to zero, you don't have a lot of information here, probably as a result of colinearity 
eigen(r)

```

**Thoughts after running this model:**

* Deciduous tree species richness was found to have an effect, but we aren't really looking at whether they use conifer forests as well. I think it would be beneficial to go back to the submodel A1 and add in a metric for native conifer species richness. 