---
title: "Full Occupancy Model 1"
author: "Anna Kurtin"
date: "2024-08-01"
output: html_document
---

This is the first iteration of the full occupacy model. From each of the submodels, I carried forward any covariate with f statistic > .70 and with a reasonable standard error. 

**Model structure:**

State model with six covariates:

* a1: tree species richness on point scale (from submodel A2)

* a2: percent canopy on landscape scale (from submodel B2) 

* a3: percent subcanopy on landscape scale (from submodel B2) 

* a4: percent subcanopy on core scale (from submodel C5)

* a5: subcanopy average height on core scale (from submodel C5)

* a6: residual variation in total veg st dev regressed on canopy height (from submodel C5)


$$
z_i \sim Bernoulli(\psi | \alpha_{1}, \alpha_{2}, \alpha_{3}, \alpha_{4}, \alpha_{5})
$$ 

Process model with four covariates:  

* b1: date

* b2: quadratic effect of date

* b3: average background decibels

* b4: survey effort

$$
y_{ij} \sim Bernoulli(p| \beta_{1}, \beta_{2}^2,\beta_{3}, \beta_{4})
$$

**Handling of NAs:** omission for detections, imputation for habitat covariates


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(rjags)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(beepr)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
```

```{r Read in Data}
full_dat <- read.csv("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Habitat_Model_Covariates/HabChap_DetOccCovsFull_SCALED_7-30.csv")

detections <- full_dat[,2:7]

# Have to specify the z data for a true presence if there was a 1 at any time 
z_dat <- rowSums(detections, na.rm = T)
z_dat[z_dat > 1] <- 1
z_dat[z_dat == 0] <- NA

# get the length of non-nas in the detection data
get_na <- function(x){(max(which(!is.na(x))))}
# Apply this to our data
miss <- apply(detections, 1, get_na)
```

```{R Setup JAGS Data}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z_dat,
  # Detection/nondetection data
  det_data = detections,
  # Native broadleaf tree spp richness
  cova1 = full_dat$tree_spp_rich,
  # % canopy landscape
  cova2 = full_dat$pct_can_landsc,
  # % subcanopy landscape
  cova3 = full_dat$pct_subcan_landsc,
  # % subcanopy core
  cova4 = full_dat$pct_subcan_core,
  # subcan avg height
  cova5 = full_dat$ht_subcan_core,
  # residual veg sd
  cova6 = full_dat$veg_sd_resid,
  # Date
  covb1 = full_dat[,35:40],
  # Background noise
  covb3 = full_dat[,22:27],
  # Effort 
  covb4 = full_dat[,29:34],
  n_unit = nrow(detections),
  miss = miss
)

# Set initial values 
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    a1 = rnorm(1, 0, 1),
    a2 = rnorm(1, 0, 1),
    a3 = rnorm(1, 0, 1),
    a4 = rnorm(1, 0, 1),
    a5 = rnorm(1, 0, 1),
    a6 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b1Q = rnorm(1, 0 ,1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1)
  )
}
```

```{R Sub Model A2 Structure}
cat("
  model{
        # Priors
    a0 ~ dlogis(0, 1)
    a1 ~ dlogis(0, 1)
    a2 ~ dlogis(0, 1)
    a3 ~ dlogis(0, 1)
    a4 ~ dlogis(0, 1)
    a5 ~ dlogis(0, 1)
    a6 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1)
    b1Q ~ dlogis(0, 1)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
  
    # Likelihood
    # Loop through each sampling unit
    for(i in 1:n_unit){
      # Impute missing habitat data
      cova1[i] ~ dnorm(0, 1)
      cova2[i] ~ dnorm(0, 1)
      cova3[i] ~ dnorm(0, 1)
      cova4[i] ~ dnorm(0, 1)
      cova5[i] ~ dnorm(0, 1)
      cova6[i] ~ dnorm(0, 1)
  
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 + a1*cova1[i] + a2*cova2[i] + a3*cova3[i] + a4*cova4[i] + a5*cova5[i] + a6*cova6[i]
      
      for(j in 1:miss[i]){
       # Impute missing detection data
        covb1[i,j] ~ dnorm(0, 1)
        covb3[i,j] ~ dnorm(0, 1)
        covb4[i,j] ~ dnorm(0, 1)
  
        det_data[i,j] ~  dbern(Z[i]*theta[i,j])  # Detection data drawn from a bernoulli distribution incorporating the true presence (Z) with probability of detection theta
        logit(theta[i,j]) <- b0 + b1*covb1[i,j] + b1Q*(covb1[i,j]^2) + b3*covb3[i,j] + b4*covb4[i,j]
        
      }
    }
    
    # Derived parameters
    
  }
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Full_Occupancy_Model/Model_Structure_Files/JAGS_GlobMod1.txt")

```

```{R Run Model, results = FALSE}
fit_g1 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Habitat_Chapter/Full_Occupancy_Model/Model_Structure_Files/JAGS_GlobMod1.txt",
                 parameters.to.save = c("a0", "a1", "a2", "a3", "a4", "a5", "a6", "b0", "b1", "b1Q", "b3", "b4"),
                 n.iter = 10000, 
                 n.burnin = 4000, 
                 n.chains = 3,
                 n.thin = 2,
                 inits = init_func)
beep(sound = 3)
```

```{R}
summary(fit_g1)
tracedens_jags(fit_g1, parmfrow = c(3,3))
```

* cova1: Native broadleaf tree spp richness

* cova2: % canopy landscape

* cova3: % subcanopy landscape

* cova4: % subcanopy core

* cova5: subcan avg height

* cova6: residual veg sd

```{R Create Violin Plots}
chains_g1 <- jags_df(fit_g1)
# Select the chains for our covariates of interest
chains_viol <- chains_g1 %>% select(a1,a2, a3, a4, a5, a6)
# Rename them to be more interpretable
colnames(chains_viol) <- c("tree spp richness", "% can landsc", "% subcan landsc", "% subcan core", "subcan avg height", "residual veg sd")
# Pivot this longer so that we can visualize it
chains_viol_long <- chains_viol %>% pivot_longer(cols = c("tree spp richness", "% can landsc", "% subcan landsc", "% subcan core", "subcan avg height", "residual veg sd"),names_to = "Parameter", values_to = "Values")

# Create the violin plot
ggplot() +
  geom_violin(data=chains_viol_long, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Point Level Covariates") +
  scale_fill_manual(values = c("tree spp richness"=palette_8[1], "% can landsc" = palette_8[2], "% subcan landsc" = palette_8[3], "% subcan core" = palette_8[4], "subcan avg height" = palette_8[5], "residual veg sd" = palette_8[6])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()

chains_beta <- chains_g1 %>% select(b1, b1Q, b3, b4)
colnames(chains_beta) <- c("date", "quadratic date", "background dB", "effort")
chains_beta_l <- chains_beta %>% pivot_longer(cols = c("date", "quadratic date", "background dB", "effort"),names_to = "Parameter", values_to = "Values")
# Create the violin plot
ggplot() +
  geom_violin(data=chains_beta_l, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Detection Covariates") +
  scale_fill_manual(values = c("date" = palette_5[1], "quadratic date" = palette_5[2], "background dB" = palette_5[3], "effort" = palette_5[4])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()
```

Questions after seeing these results:

* I think we have too many parameters in our model, but the model is converging and we're getting very significant effects even with our ten covariates. 

* What are the indications that there are too many parameters in the model?

* How large of a SE is too large of a SE in the posterior?

* What about coefficients that aren't significant anymore

* Drawing inference from this - need to back transform the parameters

* How to incorporate goodness of fit into these?