---
title: "Playback Detection Model 1"
author: "Anna Kurtin"
date: "2024-08-18"
output: html_document
---

This is the second version of the detection submodel for the playback data. I'm removing all covariates except for date so that I minimize the number of covariates in the model. 

*Effective sample size for this data* 33 * 3 = 66. So we can have a maximum of about 5-6 parameters to fit the model. 

Naive state model:

$$
z_i \sim Bernoulli(\psi_i)
$$ 

$$
logit(\psi_i) = \alpha_0
$$ 
Detection model with six covariates:  

* b1: date

* b2: quadratic effect of date

* b3: number of surveyors

* b4: start time of survey

* b5: wind condition

* b6: starting temperature



$$
y_{ij} \sim Bernoulli(p_{ij} * z_i)
$$

$$
logit(p_{ij}) = \beta_0 + \beta_{1}*date + \beta_{2}^2*date + \beta_{3}*num obs + \beta_{4}*time + \beta_5*wind + \beta_6*temp
$$

**Handling of NAs:** omission for detections, imputation for habitat covariates

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jagsUI)
library(rjags)
library(tidyverse)
library(jagshelper)
library(ggplot2)
library(beepr)
source("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/5_Visualization/Create_HexCodes_CuckooColorBlind.R")
```

```{r Read in Data}
pb_dat <- read.csv("C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/Data/Playback_Results/2023/Outputs/2023_PBData_FormatforOccModSCALED_8-6-24.csv")

detections <- pb_dat[,2:4]

# Have to specify the z data for a true presence if there was a 1 at any time 
z_dat <- rowSums(detections, na.rm = T)
z_dat[z_dat > 1] <- 1
z_dat[z_dat == 0] <- NA

# get the length of non-nas in the detection data
get_na <- function(x){(max(which(!is.na(x))))}
# Apply this to our data
miss <- apply(detections, 1, get_na)
```


```{R Look at Colinearity}
dates <- pb_dat %>% select(site_id, date_s1, date_s2, date_s3)

# reference code for making the columns long and then combining them
# dates_long <- dates %>% pivot_longer(cols = c(start_date_s1,
#                                               start_date_s2,
#                                               start_date_s3,
#                                               start_date_s4,
#                                               start_date_s5,
#                                               start_date_s6), names_to = "survey_period",values_to = "date" )
# # rename to s1, s2, etc
# dates_long <- dates_long %>% mutate(survey_period = case_when(survey_period == "start_date_s1" ~ "s1",
#                                                 survey_period == "start_date_s2" ~ "s2",
#                                                 survey_period == "start_date_s3" ~ "s3",
#                                                 survey_period == "start_date_s4" ~ "s4",
#                                                 survey_period == "start_date_s5" ~ "s5",
#                                                 survey_period == "start_date_s6" ~ "s6"))
# # combine
# dates_long <- unite(dates_long,site_survey,c(site_id, survey_period),sep="_",remove=TRUE)

obs <- pb_dat %>% select(site_id, num_obs_s1, num_obs_s2, num_obs_s3)
time <- pb_dat %>% select(site_id, time_s1, time_s2, time_s3)
wind <- pb_dat %>% select(site_id, wind_s1, wind_s2, wind_s3)
temp <- pb_dat %>% select(site_id, temp_s1, temp_s2, temp_s3)
```

```{R Setup JAGS Data}
# Create a list of data for JAGS
jags_data <- list(
  # True presence at sites with confirmed presences
  Z = z_dat,
  # Detection/nondetection data
  det_data = detections,
  # Date
  covb1 = dates[,2:4],
  # num surveyors
  covb3 = obs[,2:4],
  # start time 
  covb4 = time[,2:4],
  # wind condition
  covb5 = wind[,2:4],
  # temp
  covb6 = temp[,2:4],
  n_unit = nrow(detections),
  miss = miss
)

# Set initial values 
init_func <- function(){
  list(
    a0 = rnorm(1, 0, 1),
    b0 = rnorm(1, 0, 1),
    b1 = rnorm(1, 0, 1),
    b1Q = rnorm(1, 0 ,1),
    b3 = rnorm(1, 0, 1),
    b4 = rnorm(1, 0, 1),
    b5 = rnorm(1, 0, 1),
    b6 = rnorm(1, 0 ,1)
  )
}
```

```{R PB Model 1 Structure, eval = FALSE}
cat("
  model{
        # Priors
    a0 ~ dlogis(0, 1)
    b0 ~ dlogis(0, 1)
    b1 ~ dlogis(0, 1)
    b1Q ~ dlogis(0, 1)
    b3 ~ dlogis(0, 1)
    b4 ~ dlogis(0, 1)
    b5 ~ dlogis(0, 1)
    b6 ~ dlogis(0, 1)
  
    # Likelihood
    # Loop through each sampling unit
    for(i in 1:n_unit){
  
      Z[i] ~ dbern(psi[i])  # Presence or absence drawn from a bernoulli distribution with occupancy probability psi  
      logit(psi[i]) <- a0 
  
      for(j in 1:miss[i]){
       # Impute missing detection data
        covb1[i,j] ~ dnorm(0, 1)
        covb3[i,j] ~ dnorm(0, 1)
        covb4[i,j] ~ dnorm(0, 1)
        covb5[i,j] ~ dnorm(0, 1)
        covb6[i,j] ~ dnorm(0, 1)
  
        det_data[i,j] ~  dbern(mu_p[i,j])  # Create probability density drawn for each site
        mu_p[i,j] <- Z[i]*theta[i,j]
        logit(theta[i,j]) <- p[i,j]
        p[i,j] <- b0 + b1*covb1[i,j] + b1Q*(covb1[i,j]^2) + b3*covb3[i,j] + b4*covb4[i,j] + b5*covb5[i,j] + b6*covb6[i,j]
        # Det NA
        det_na[i,j] <- det_data[i,j]
  
        #Calculate Bayes p-value - plug the parameters into the equation for each sample of the posterior and use this to create new observations
        Presi[i,j] <- (det_na[i,j] - theta[i,j])^2 # Calculate squared residual of observed data
        y_new[i,j] ~ dbern(mu_p[i,j])             # Simulate observed data
        Presi_new[i,j] <- (y_new[i,j] - theta[i,j])^2 # Calculate squared residual error of simulated data
      }
    }
    
    # Derived parameters
    ## Calculate sum of squared residual errors for observed 
    for(i in 1:n_unit){
       sobs[i] <- sum(Presi[i,1:miss[i]])
       ssim[i] <- sum(Presi_new[i,1:miss[i]])
    }
    
    SSEobs <- sum(sobs[])
    SSEsim <- sum(ssim[])
    p_val <- step(SSEsim - SSEobs)  
    
  }
  ", file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Methods_Chapter/Playback_Model/Model_Structure_Files/JAGS_PBMod1.txt")

```

```{R Run Model, results = FALSE}
fit_1 <- jags(data = jags_data, 
                 model.file = "C:/Users/annak/OneDrive/Documents/UM/Research/Coding_Workspace/Cuckoo-Research/R_Scripts/2_Analyses/Methods_Chapter/Playback_Model/Model_Structure_Files/JAGS_PBMod1.txt",
                 parameters.to.save = c("a0", "b0", "b1", "b1Q", "b3", "b4", "b5", "b6", "p_val"),
                 n.iter = 10000, 
                 n.burnin = 5000, 
                 n.chains = 3,
                 n.thin = 2,
                 inits = init_func)
beep(sound = 3)
```

```{R}
summary(fit_1)
tracedens_jags(fit_1, parmfrow = c(3,3))
```


```{R Create Violin Plots}
chains_1 <- jags_df(fit_1)

chains_beta <- chains_1 %>% select(b1, b1Q, b3, b4, b5, b6)
colnames(chains_beta) <- c("date", "quadratic date", "number_obs", "start_time", "wind_strength","temp")
chains_beta_l <- chains_beta %>% pivot_longer(cols = c("date", "quadratic date", "number_obs", "start_time", "wind_strength","temp"),names_to = "Parameter", values_to = "Values")
# Create the violin plot
ggplot() +
  geom_violin(data=chains_beta_l, aes(x = Parameter, y = Values,fill = Parameter),trim = FALSE) +
  labs(title = "Posterior Distribution of Detection Covariates") +
  scale_fill_manual(values = c("date" = palette_8[1], "quadratic date" = palette_8[2], "number_obs" = palette_8[3], "start_time" = palette_8[4], "wind_strength" = palette_8[5], "temp" = palette_8[6])) + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  theme_minimal()
```